---
title: "Finding probabilities"
author: "SMWadgymar"
date: '2022-08-12'
output: html_document
---


The binomial distribution is the distribution that describes the frequency of outcomes for each set of trials where there are exactly two possible outcomes for each trial (e.g. heads or tails, survives or dies, etc.).  The notation used to describe the binomial distribution, B(n,p), includes two parameters: n is the size (# flips) and p is the probability that each is heads (for now, 0.5).  A variable is binomially distributed if (1) the trials are independent, (2) the number of trials is fixed, (3) each trial outcome can be classified as success or failure, and (4) the probability of success is the same for each trial. This is one of many distributions and we will touch on a few distributions (not all by any means!) in class. 

R has many built-in functions that allow us to draw samples from well-characterized distributions.  We’ll make use of three of those functions for the binomial distribution here: ‘rbinom’, ‘dbinom’, and ‘pbinom’. Importantly, we will make use of these functions to examine probabilities in two ways. First, we’ll run simulations, as we did before. Simulations are digital experiments. We use them to conduct large, replicated experiments that would not be possible in real life, and as in real life, we want our outcomes to be the product of random chance. The goal is to run enough simulations that the probabilities that we estimate are likely very close approximations of the true probabilities (which are often unknown). For our second approach, we will calculate exact probabilities. We can do this because we are drawing our samples from a well-characterized distribution where the exact probabilities are known.


####Finding probabilities through simulation

6) Simulate 100000 draws each with 10 fair coins and save the outcome as 'flips'.

```{r}
flips<-rbinom(???, ???, ???)
```

Q: How many values are stored in the variable called flips?  What is your interpretation of each value?

A: 


7) Let's find the fraction of draws where 5 heads occurred.

```{r}
mean(flips==5)
```

Q: Which do you expect will be lower: the probability that 3 out of 10 flips will be heads or the probability that 5 out of 10 flips will be heads?  Why?  

A: 


8) Next, calculate the probability that 3 out of 10 flips will be heads.  This reflects the probability that this outcome will occur, and is called the density of this distribution at the point where x=3.

```{r}

```

Q: How does the probability that 3/10 flips will be heads compare to the probability that 5/10 flips will be heads?

A:


9) Let's find the cumulative probability that heads will occur 4 or fewer times during each draw.  This is referred to as the cumulative probability because it sums up the probability that the number of heads is 0, 1, 2, 3, or 4 out of 10 flips.

```{r}
mean(flips<=4)
```

Q: How would you adjust this line of code to estimate the cumulative probability that heads will occur 6 or more times?

A:
```{r}

```



####Finding the exact probabilities

10) The dbinom function estimates the exact probability density at a given point.  The arguments here are the density being estimated, the number of coins, and the probability of producing a head.  Note that the order and type of arguments used here do not match the order and type of arguments that you used for the 'rbinom' function.

```{r}
dbinom(5, 10, 0.5)
```

Q: How does this command compare to the steps you took for question 6 and 7 above?  Are the estimates comparable?  Why or why not?

A: 


11) Now let's estimate the exact cumulative probability density for a given range.  The arguments here are the density being estimated, the number of coins, and the probability of producing a head.  Note that the order and type of arguments used here do not match the order and type of arguments that you used for the 'rbinom' function.

```{r}
pbinom(4, 10, 0.5)
```

Q: How does this command compare to the steps you took for question 9?  Are the estimates comparable?  Why or why not?

A: 




PRACTICE: Do not use ChatGPT to answer these questions - try them on your own based on the information above. For each of the following scenarios, identify which function (rbinom, dbinom, or pbinom) you would use to find the solution.

Scenario 1: You are studying the effect of a toxin on a frog population. It is known that 40% of the frogs exposed to this toxin will die. You want to simulate how many frogs out of a sample of 100 survive after being exposed to the toxin.

A: 

Scenario 2: A predator has a 30% chance of catching its prey during a hunt. You want to find the cumulative probability that the predator will catch 3 or fewer prey in 10 hunts.

A:

Scenario 3: In a population of 1000 birds, 5% are infected with a virus. You want to simulate the number of infected birds in a random sample of 100 birds.

A: 

Scenario 4: In a population of plants, there is a 10% chance that any given offspring will carry a certain genetic mutation. You have 50 offspring and want to calculate the probability of getting exactly 5 mutants in this population.

A: 

Scenario 5: There is a 20% chance of detecting an endangered species in a specific region during a survey. You want to find the probability that in 15 surveys, the species is detected at least once.

A: 

Scenario 6: You are testing whether light exposure affects the blooming of flowers. You know from previous studies that 70% of flowers bloom when exposed to sunlight. In an experiment with 20 flowers, you want to calculate the probability that 14 flowers will bloom.

A: 

Scenario 7: In a fish population, 15% are infected by a certain parasite. You want to calculate the probability of observing 8 or fewer infected fish in a sample of 50.

A: 

Scenario 8: A bird species has a 60% chance of successfully hatching eggs. You are interested in simulating how many out of 5 eggs in a nest will hatch.

A: 

Scenario 9: In a population of insects, 45% of offspring exhibit a specific trait. You want to simulate the number of offspring with this trait in 50 random insect births.

A: 

Scenario 10: You are studying the expression of a gene and know that each gene copy has a 25% chance of being expressed. In a group of 12 gene copies, you want to determine the probability that 3 of them will be expressed.

A: 
