---
title: 'Normal data: How much data is within 1 SD of the mean?'
author: "SMWadgymar, KGSmith"
date: '2022-08-11'
output: html_document
---

This script will guide you through doing some simple calculations in R to test "the empirical rule". 

This rule of thumb states that,for normal distributions,:

about 68% of the data will be within ±1 standard deviation (sd) of the mean value

about 95% of the data will be within ±2 sd of the mean.

Here, we will be exploring how the shape of a dataset affects this rule and our interpretation of sd.



####Normal data

First, we will create a dataset drawn randomly from a normal distribution.  Doing this will ensure that the data we are working with are pretty much perfectly normal, which means this is a theoretical test of the empirical rule.

Here is how we will simulate the data:

```{r}
norm.data=rnorm(10000)
```

The above creates a dataset of 10,000 values drawn from a noraml distribution with a mean of 0 and an sd of 1 (the default settings in R).

Let's first explore the shape of the dataset to confirm that it's normal.  Use words other than 'normal' to describe the shape of the histogram you expect to see.

```{r}
hist(???, col="darkolivegreen4", breaks=50, xlab="# of sd from the mean")
```

Below, print the mean and SD of this data using mean() and sd().

```{r}

```

Q: Based on the distribution of the data, do you think the empirical rule will apply and that ~68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 



We can actually find out how many datapoints are within +/- 1 sd of the mean using a few simple steps.

1) You know the mean and sd of the data.  From these values, define the boundaries of an area that is 1 sd wide on either side of the mean value.  

```{r}
lower = mean(???) - sd(???)

upper = ???
```

2) Using these boundaries, create a subset that includes only the data in the middle of the range (i.e., +/- 1 sd from the mean).  There are several ways to do this, but here is one way that uses the subset() function.

```{r}
middle = subset(norm.data, norm.data <= ??? & norm.data >= ???)
```


3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all data is within +/- 1 sd of the mean.

```{r}

```

Q: What proportion of data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this datset?  Why do you think that is?

A: 




####Human height data

First, read the data in.

```{r}
source("http://www.openintro.org/stat/data/cdc.R")
```

Use the code from the first dataset above to answer the following questions about the human height dataset.

Look at the distribution of this data using a histogram.

```{r}

```

Find the mean and standard deviation of height using mean() and sd()

```{r}

```

Q: Do you think the "empirical rule" will apply and that 68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 


Find out how many datapoints are within +/- 1 sd of the mean in three steps.

1) Define the boundaries of a zone that is 1 sd wide on either side of the mean value of height.

```{r}

```

2) Using these boundaries, create a subset that includes only the data in the middle of this range (i.e., +/- 1 sd from the mean).

```{r}

```

3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all the data is within +/- 1 sd of the mean.

```{r}

```

Q: What proportion of the data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this dataset?  Why do you think this is?

A:




##Transcript length data

This dataset contains transcript length data.

First, read the data in.

```{r}
coding.mrna=read.csv("https://docs.google.com/spreadsheets/d/1gnZefB6AQ2hSoZaSp4r40cxQ7vi4KIKnGDNkdCgAjKc/gviz/tq?tqx=out:csv&sheet=coding.mrna")
```


Let's create a datset that includes only the trascript lengtsh from Chromosome 2, which will make things a bit easier.  There are many different ways to subset data.  

```{r}
chr2.length=subset(coding.mrna, ???)
```


Use the code from the dataset above to answer the following questions about the transcript length data.

Look at the distribution of transcript length using a histogram.

```{r}

```

Find the mean and standard deviation of transcript length using mean() and sd()

```{r}

```

Q: Do you think the "empirical rule" will apply and that 68% of data values will be within +/- 1 sd of the mean?  Why or why not?

A: 


Find out how many datapoints are within +/- 1 sd of the mean in three steps.

1) Define the boundaries of a zone that is 1 sd wide on either side of the mean value of height.

```{r}

```

2) Using these boundaries, create a subset that includes only the data in the middle of this range (i.e., +/- 1 sd from the mean).

```{r}

```

3) Lastly, find out how many data values are within this middle range and then calculate what proportion of all the data is within +/- 1 sd of the mean.

```{r}

```

Q: What proportion of the data fell within +/- 1 sd of the mean?  What does this mean with respect to how well the empirical rule worked for this dataset?  Why do you think this is?

A:


